---
title: "prettydates"
author: "ebinabo"
date: "December 28, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the documentation for the package at <https://github.com/ebinabo/prettydates>

A beginner level knowledge of R is assumed.

Totally new to R? Some resources are cited at the bottom of the page.

Package can be installed using `install_github('ebinabo/prettydates')`

## Load the Libaries

```{r libraries}
library(prettydates)
library(purrr) # map functions
library(knitr) # to use the kable function for pretty tables
```

The library takes in dates in `Date`, `POSIXct`, `POSIXt`, etc formats and prints them to the screen in a more easily readable format.

These are particularly useful for creating documents where you might want the date out in full, so they could be easily read out or skimmed through.

Let's list out all the functions currently listed in the package...

Some of the prefixes and suffixes used and their descriptions for a better overview;

- `long_` dates with weekday, month, year in full
- `short_` dates with weekday, month, year in short
- `_day_month_` displaying day and month, w/o year
- `_dt12` datetime in a 12hr format
- `_dt24` datetime in a 24hr format
- `_w` dates including day of week
- `_us` dates in the american format

```{r ls functions, echo=FALSE}
unclass(lsf.str(envir = asNamespace("prettydates"), all = T))
```

## Usage

Functions can be used on different classes of objects including;

- a single date or datetime object
- a vector of dates or datetime objects
- a column of dates or datetime objects
- a column of dates or datetime objects within a list using the map() family of functions

Here are some examples

### Single Date or Datetime object

```{r single}
long_date(Sys.Date())
long_dt12_us_w(Sys.time())
```


### Vector of Dates or Datetime objects

```{r vector}
seq_date <- seq.Date(Sys.Date()-70, Sys.Date(), 7)
short_date(seq_date)

times <- c(Sys.time(), Sys.time()-1e02, Sys.time()-1e03, Sys.time()-1e04, Sys.time()-1e05)
short_dt24_us(times)
```

### Column of Dates or Datetime objects

Lets create a sample dataframe showing different orders made on different days

```{r df init}
order_id <- paste("Order", 1:10, sep = " ")
order_dates <- seq.Date(Sys.Date()-70, Sys.Date(), 1)
order_dates <- sample(order_dates, 10)
order_history <- as.data.frame(order_dates)
row.names(order_history) <- order_id
names(order_history) <- "Date"
kable(order_history)
```

Now let's use one of the funtions from prettydates to change the display rendered. The time of order is set to a default 12:00:00 AM because we're using a sample of dates

```{r df w prettydates}
order_history$Date <- short_dt12_us_w(order_history$Date)
kable(order_history)
```

### Column of Dates or Datetime objects in a list

For more advanced users, the current block of code doesn't return any errors but the desired output which should look like "28 December" returns a different output

If you have a clue how to fix this, please fork me on github <https://github.com/ebinabo>

```{r}
order_history_1 <- order_history
order_history_1$Date <- map(order_history_1$Date, long_day_month)
kable(order_history_1)
```

### All the functions in action

Here are what all the functions should return if used on the current date

```{r in action, error=T}
todays_date <- Sys.Date()
current_time <- Sys.time()

long_date(todays_date) #1
long_date_us(todays_date) #2
long_date_us_w(todays_date) #3
long_date_w(todays_date) #4
long_day_month(todays_date) #5
long_day_month_dt12(current_time) #6
long_day_month_dt12_us(current_time) #7
long_day_month_dt12_us_w(current_time) #8
long_day_month_dt12_w(current_time) #9
long_day_month_dt24(current_time) #10
long_day_month_dt24_us(current_time) #11
long_day_month_dt24_us_w(current_time) #12
long_day_month_us(todays_date) #13
long_day_month_us_w(todays_date) #14
long_day_month_w(current_time) #15
long_dt12(current_time) #16
long_dt12_us(current_time) #17
long_dt12_us_w(current_time) #18
long_dt12_w(current_time) #19
long_dt24(current_time) #20
long_dt24_us(current_time) #21
long_dt24_us_w(current_time) #22
long_dt24_w(current_time) #23

short_date(todays_date) #24
short_date_us(todays_date) #25
short_date_us_w(todays_date) #26
short_date_w(todays_date) #27
short_day_month(todays_date) #28
short_day_month_dt12(current_time) #29
short_day_month_dt12_us(current_time) #30
short_day_month_dt12_us_w(current_time) #31
short_day_month_dt12_w(current_time) #32
short_day_month_dt24(current_time) #33
short_day_month_dt24_us(current_time) #34
short_day_month_dt24_us_w(current_time) #35
short_day_month_us(todays_date) #36
short_day_month_us_w(todays_date) #37
short_day_month_w(current_time) #38
short_dt12(current_time) #39
short_dt12_us(current_time) #40
short_dt12_us_w(current_time) #41
short_dt12_w(current_time) #42
short_dt24(current_time) #43
short_dt24_us(current_time) #44
short_dt24_us_w(current_time) #45
short_dt24_w(current_time) #46
```


### Resources

Some useful resources are cited below to help you grasp some of these functions better

<https://www.datacamp.com/courses/free-introduction-to-r> for an introduction to R

<https://jennybc.github.io/purrr-tutorial> for a purrr tutorial

<https://www.datacamp.com/courses/intermediate-functional-programming-with-purrr> for creating mappers


